\pkg{cleanR}: Maid for Cleaning Datasets in \proglang{R}

### 2016-12-16 latex/article_vol2.tex
Proof-reading, minor stuff, structure.

%% an abstract and keywords
\Abstract{Data cleaning and -validation are the first steps in any data
  analysis, as the validity of the conclusions from the analysis are
  hinged on the quality of the input data. Mistakes in the data can arise for any number of reasons, including erroneous codings, malfunctioning measurement equipment, inconsistent data generation manuals and many more.  Ideally, a human investigator
  should go through each variable in the dataset and look for
  potential errors --- both in input values and coding.  \hl{segway}

  We describe an \proglang{R} package which implements an extensive
  and customizeable suite of checks to be applied to the variables in
  a dataset in order to identify potential problems in the
  corresponding variables. The results can be presented in an auto-generated, non-technical, stand-alone
  overview document, intended to be perused by an investigator with an
  understanding of the variables in the data and the experimental design, but not necessarily knowledge of \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysist and field experts, while also providing easy documentation of data cleaning steps and data quality control. Moreover, the \R{cleanR} solution changes the data cleaning process from the usual ad hoc approach to a systematic, well-documented endeavor. 
  \R{cleanR} also provides a suite of more typical \R{R} tools for interactive data quality assessment and -cleaning.

  The \pkg{cleanR} package is designed to be easily extended with custom
  user-created checks that are relevant in particular situations.
}
\Keywords{data cleaning, quality control, \proglang{R}}
\Plainkeywords{data cleaning, quality control, R} %% without formatting
%% at least one keyword must be supplied

@@

\section{Introduction}
Statisticians and data analysts spend a large portion of their time on
data cleaning and on data wrangling. Packages such as
\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier in
\proglang{R}, but there are only a few tools available for data cleaning. \hl{What are these tools? More here about the solutions already available.}

\hl{A text here (with offset in what tools are already available) that
\begin{itemize}
\item Makes the point that data cleaning is usually done in a very ad hoc manner
\item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
\item Gives a specification of what we mean by data errors/mistakes/whatever. I'm not completely sure this will be clear to everyone.
\end{itemize}
}

Data cleaning is a time consuming endeavor, as it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of their origin. 
\hl{Segway: Something like maybe this is why proper data cleaning is not always done...}
In many situations these errors are discovered in the process of the
data analysis (e.g., \hl{a categorical variable with numeric labels is
wrongly classified as a numeric variable\footnote{Is that something we can fix in clean()?}} or a variable where all values have erroneously been coded to the same value), but in other cases a human
with knowledge about the data context area is needed to identify
possible mistakes in the data (e.g., if there are 4 categories for a
variable that should only have 3).  \hl{segway} \\

The \R{cleanR} approach to data cleaning and -quality assessment is characterized by two principles. First of all, there is no need for data cleaning to be an ad hoc procedure. Often, we have a very clear idea of what flags are raisable in a given dataset before we look at it, as we were the ones to produce it in the first place. This means that data cleaning can easily be a well-documented, well-specified procedure. This angle on data cleaning also adds to the transparency of this first step of data analysis: Though data cleaning might be regarded as a somewhat dull activity, adequate data cleaning is crucial in any data analysis. In order to aid this principle, \R{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build. This quality assessment is presented in a auto-generated overview document, readable by data analysts and field experts alike, thereby also contributing to a inter-field dialogue about the data at hand. Oftentimes, e.g. separating faulty codings of a numeric value from unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst. Hopefully, having easy access to data descriptions through \R{cleanR} will help this necessary knowledge sharing.

While \R{cleanR} is build for auto-generating data quality assessment overview documents, we also wish to emphasize that it is \textit{not} a tool for automatic data cleaning. This qualifies as the second principle of \R{cleanR}: Data cleaning decisions should always be made by humans. Therefore, \R{cleanR} does not provide any tools for "fixing" errors in the data. However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues.\\

This manuscript is structured as follows: First, we introduce the representative of the first principle, namely the \R{clean()} function, which generates data cleaning overview documents. In the \R{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common. However, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions. In Section 3 \hl{proper reference here}, we turn to the question of how such \R{cleanR} extensions can be made such that they are integrable with the \R{clean()} function and with other tools available in \R{cleanR}. In this section, we also present the interactive mode of \R{cleanR}, as motivated by the second principle above. At last, in Section 4 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \R{cleanR} can be used to solve them.

@@

\begin{figure}
\begin{center}
\frame{\includegraphics[width=7.5cm,page=2]{cleanR_toyData.pdf}} 
\frame{\includegraphics[width=7.5cm, page=3]{cleanR_toyData.pdf}}
%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
\end{center}
\label{fig:example1}
\caption{Example output from running \R{clean()} on the \R{toyData} dataset. First a summary of the full dataset,.. 
  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small.}}
\end{figure}

\begin{figure}
\begin{description}
\item[Input] A dataset
	\begin{itemize} 
		\item This should be of type \R{data.frame} or \R{tibble}
	\end{itemize}
\item[Create contents] For each variable in the dataset, do the following:
	\begin{description}
		\item[Stage 1:] Pre-checks 
			\begin{itemize}
				\item Is the variable suitable for summarization, visualization and checks?
				\begin{description}
					\item[Yes:] Go to stage 2 
					\item[No:] Move on to the next variable
				\end{description}
			\end{itemize}
		\item[Stage 2:] SVC-steps
			\begin{description}
				\item[Summarize] Call \R{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
				\item[Visualize]  Call \R{visualize()} to produce a plot visualizing the distribution of the variable. 
				\item[Check]  Call \R{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
			\end{description}
	\end{description}
\item[Output] Files for a overview document 
	\begin{itemize}
		\item Always a \R{rmarkdown} (.Rmd) file
		\item Possible also a html or pdf file
	\end{itemize}
\end{description}
\caption{\hl{quite odd having a figure like this? maybe do a flowchart?}}
\label{figure:cleanStructure}
\end{figure}

In \pkg{cleanR}, the \code{clean} function is the primary workhorse and knowledge of nothing more than this function is required to produce data cleaning outputs. The data cleaning output itself is an overview document, intended for reading by humans, in either pdf or html format. Appendix \hl{something} provides an example of a data cleaning output document, produced by calling \R{clean} on the dataset \R{toyData} available in  \R{cleanR}. The first two pages of this data cleaning output are also available in Figure~\ref{fig:example1}. \R{toyData} is a very small ($15$ by $6$), artificial dataset, whose only purpose is to illustrate the main capabilities of \R{cleanR}. The following commands load the dataset and produce the cleaning output:

@@

By default, a pdf overview document is produced, saved to the disc (in the working directory) and opened for inspection. Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts. First, an overview of what was done is presented under the title \textit{Data cleaning summary}. Secondly, each variable in the dataset is presented in turn using (up to) three tools in the \textit{Variable list}: A table summarizing key features of the variable, a figure visualizing its distribution and potentially also a list of flagged issues. For instance, in the \R{numeric}-type variable \R{var2} from \R{toyData}, \R{clean()} has identified two values that are suspected to be miscoded missing values (\R{999} and \R{NaN}), while two values were also flagged as potential outliers that should be investigated more carefully. We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable. \hl{hm, delete the last part? or use different variable? Only these exact two checks were performed}. \\

\hl{metatext here: blabla more can be done blabla...} \\

@@

\begin{table}
\small
\begin{tabular}{p{0.2\linewidth}p{0.5\linewidth}p{0.2\linewidth}}
\hline
Argument & Description & Default value \\
\hline

\smallskip Control \hl{...eh?}\\
\quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
\quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the variable list? & \R{FALSE} \\
\quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\

@@

\smallskip Control output \\
\quad \R{output} & Type of output file to be produced (html, pdf or \hl{?? what does screen mean here??}) & \R{"pdf"} \\
\quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards? & \R{TRUE} \\

\hline
\end{tabular}
\caption{A selection of commonly used arguments of \R{clean()}.  \hl{Note: I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions (dealt with in next subsection)}.  \hl{Fix formatting for \R{onlyProblematic}.}}
\label{table.cleanFormals}
\end{table}

\hl{Describe arguments in the table...}

@@

\begin{table}
\centering
\begin{tabular}{p{0.3\linewidth} p{0.35\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
 p{0.01\linewidth} p{0.01\linewidth}}
  \hline
& Description &  \multicolumn{7}{c}{Variable classes} \\ \smallskip
 & &  C & F & I & L & B & N & D\\ 
  \hline \smallskip
  \textbf{\R{summaryFunction}s}  \smallskip \\
  \quad \R{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\ 
  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ &  \\ 
  \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \smallskip \\
 \textbf{\R{visualFunction}s} \smallskip \\
  \quad \R{basicVisual} & Histograms and barplots using graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \smallskip \\
 \textbf{\R{checkFunction}s} \smallskip \\
 \quad \R{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\ 
 \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & \\ 
 \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ &  \\ 
 \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\ 
 \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & \smallskip   \\ 
 \hline
\end{tabular}
\caption{\hl{Blabla, mention that C is character, F is factor, I is integer, L is labelled, B is logical (boolean), N is numeric and D is Date. \newline Also: Check that everything in here is correct (i.e. corresponds to the output of allSummaryFunctions() etc}}
\label{table.SVCfunctions}
\end{table}

\hl{
\begin{itemize}
\item Introduce the relevant \R{clean}-arguments and the \R{defaultWhateverSummaries} etc.-functions
\item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
\item Small example
\end{itemize}
}

By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset. There are two stages in which this occurs \hl{(with ref. to figure/flowchart?)}:
\begin{enumerate}
\item In the precheck functions
\item In the summarize/visualize/check (SVC) step
\end{enumerate}
Each of these stages are controllable using appropriate function parameters in \code{clean}. In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs. However, if for instance the dataset at hand requires completely different visualizations, more control is needed. \pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s. They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes. By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands. Therefore, this section is dedicated to introducing each in turn.\\

\section{Something about the interactive cleanR mode}
While overview documents are great for documenting our work, they might not feel very natural for actually performing data cleaning or data wrangling based on the results we found. Therefore, \R{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use. Or rather, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively. \hl{segway}

\hl{Maybe write here that this section introduces the three functions: \R{summarize()}, \R{check()} and \R{visualize()}?}

\subsection{Blabla: An example}
Let's say we wish to look further into a certain variable, namely \R{var2}, from \R{toyData}. The data cleaning summary found some issues, and we would like to recall what these issues were. This is done using the command
\begin{Schunk} 
\begin{Sinput} 
> check(toyData$var2)
\end{Sinput}
\begin{Soutput}
$identifyMissing
The following suspected missing value codes enter as regular values: 999, NaN.
$identifyOutliers
Note that the following possible outlier values were detected: 82, 999. $
\end{Soutput}
\end{Schunk}
\hl{delete that last \$, it's only there to make my TeX editor stop highlighting everything. Note: \$s cannot be escaped here, as Schunk is verbatim.}

@@

\begin{Schunk}
\begin{Sinput}
> missVar2 <- identifyMissing(toyData$var2)
> str(missVar2)
\end{Sinput}
\begin{Soutput}
List of 3
 $ problem      : logi TRUE
 $ message      : chr "The following suspected missing value codes enter as 
 		 regular values: \\\"999\\\", \\\"NaN\\\"."
 $ problemValues: num [1:2] 999 NaN
 - attr(*, "class")= chr "checkResult"
\end{Soutput}
\end{Schunk}
The most important thing to note here is that while the printed message is made for human reading, the actual values of the variable causing the issue are still attainable. If we for instance decide that the values \R{999} and \R{NaN} in \R{var2} are in fact miscoded missing values, we can replace them easily:
\begin{Schunk}
\begin{Sinput}
> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
> identifyMissing(toyData$var2}
\end{Sinput}
\begin{Soutput}
No problems found. $
\end{Soutput}
\end{Schunk}
\hl{remove that last \$}. \hl{This is quite a practical example, but it essentially also gives the basic structure for misusing cleanR by generating a loop that does autocleaning... What do you think? Perhaps those who know how to make loops also know that they shouldn't do automatic data cleaning?}

\hl{
\begin{itemize}
\item  Mention allCheckFunctions() etc. again here
\item Mention check(), visualize() and summarize() modes for data.frames
\end{itemize}
}

@@

\subsection{Writing cleanR extensions: Function templates}
\hl{metatext}

>>>>

%% an abstract and keywords
\Abstract{Data cleaning and -validation are the first steps in any data
  analysis, as the validity of the conclusions from the analysis are
  hinged on the quality of the input data. Mistakes in the data can arise for any number of reasons, including erroneous codings, malfunctioning measurement equipment, inconsistent data generation manuals and many more.  Ideally, a human investigator
  should go through each variable in the dataset and look for
  potential errors --- both in input values and coding.  \hl{segway: But, ...}

  We describe an \proglang{R} package which implements an extensive
  and customizeable suite of quality assessment tools to be applied to
  a dataset in order to identify potential problems in its variables. The results can be presented in an auto-generated, non-technical, stand-alone
  overview document, intended to be perused by an investigator with an
  understanding of the variables in the data and the experimental design, but not necessarily knowledge of \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysist and field experts, while also providing easy documentation of data cleaning steps and data quality control. Moreover, the \R{cleanR} solution changes the data cleaning process from the usual ad hoc approach to a systematic, well-documented endeavor. 
  \R{cleanR} also provides a suite of more typical \R{R} tools for interactive data quality assessment and -cleaning.

  The \pkg{cleanR} package is designed to be easily extended with custom
  user-created checks that are relevant in particular situations. \hl{Already said that above. Either expand on it or delete this.}
}
\Keywords{data cleaning, quality control, \proglang{R}}
\Plainkeywords{data cleaning, quality control, R} %% without formatting
%% at least one keyword must be supplied

@@

\section{Introduction}
Though data cleaning might be regarded as a somewhat dull activity, adequate data cleaning is crucial in any data analysis. With ever-growing dataset sizes and complexities, statisticians and data analysts find themselves spending a large portion of their time on
data cleaning and on data wrangling. \proglang{R}-packages such as
\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier, but there are only a few tools available for data cleaning. \hl{What are these tools? More here about the solutions already available.}

\hl{A text here (with offset in what tools are already available) that
\begin{itemize}
\item Makes the point that data cleaning is usually done in a very ad hoc manner
\item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
\item Gives a specification of what we mean by data errors/mistakes/whatever. I'm not completely sure this will be clear to everyone and we should probably mention it early on to not lose readers/users.
\end{itemize}
}

Data cleaning is a time consuming endeavor, as it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of their origin. 
\hl{Segway: Something like maybe this is why proper data cleaning is not always done...}
In many situations these errors are discovered in the process of the
data analysis (e.g., \hl{a categorical variable with numeric labels is
wrongly classified as a numeric variable\footnote{\hl{Is that something we can fix in clean()? Not quite sure what "labels" mean in this context.}}} or a variable where all values have erroneously been coded to the same value), but in other cases a human
with knowledge about the data context area is needed to identify
possible mistakes in the data (e.g., if there are 4 categories for a
variable that should only have 3).  \hl{segway} \\

The \R{cleanR} approach to data cleaning and -quality assessment is characterized by two principles \hl{use a different term than "principles" here and below, when we refer to this sentence}. First of all, there is no need for data cleaning to be an ad hoc procedure. Often, we have a very clear idea of what flags are raisable in a given dataset before we look at it, as we were the ones to produce it in the first place. This means that data cleaning can easily be a well-documented, well-specified procedure. In order to aid this principle, \R{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build. This quality assessment is presented in a auto-generated overview document, readable by data analysts and field experts alike, thereby also contributing to a inter-field dialogue about the data at hand. Oftentimes, e.g. distinguishing between faulty codings of a numeric value and unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst. Hopefully, having easy access to data descriptions through \R{cleanR} will help this necessary knowledge sharing.

While \R{cleanR}s primary raison d'être is auto-generating data quality assessment overview documents, we still wish to emphasize that it is \textit{not} a tool for automatic data cleaning. This qualifies as the second principle of \R{cleanR}: Data cleaning decisions should always be made by humans. Therefore, \R{cleanR} does not supply any tools for "fixing" errors in the data. However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues, one variable at a time.\\

This manuscript is structured as follows: First, in section \hl{2}, we introduce the representative of the first \hl{principle}, namely the \R{clean()} function, which generates data cleaning overview documents. In the \R{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common. Next, in section \hl{3}, we present the interactive mode of \R{cleanR}, as motivated by the second \hl{principle} above. But, as any data analyst knows, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions. In Section \hl{4}, we turn to the question of how such \R{cleanR} extensions can be made, such that they are integrable with the \R{clean()} function and with the other tools available in \R{cleanR}.  At last, in Section 5 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \R{cleanR} can be used to solve them. \hl{Is there a Section 6?}\\

\hl{Do they typically discuss notation in articles like this? I try (somewhat inconsistently) to refer to functions as \R{function()} while other \R{R} objects are referred to merely as \R{object}. This is also the style adopted by Wickham in his books.}

@@

\begin{figure}
\begin{center}
\frame{\includegraphics[width=7.5cm,page=2]{cleanR_toyData.pdf}} 
\frame{\includegraphics[width=7.5cm, page=3]{cleanR_toyData.pdf}}
%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
\end{center}
\label{fig:example1}
\caption{Example output from running \R{clean()} on the \R{toyData} dataset. First a summary of the full dataset,.. 
  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small, right?}}
\end{figure}

\begin{figure}
\begin{description}
\item[Input] A dataset
	\begin{itemize} 
		\item This should be of type \R{data.frame} or \R{tibble} \hl{or data.table?}
	\end{itemize}
\item[Create contents] For each variable in the dataset, do the following:
	\begin{description}
		\item[Stage 1:] Pre-checks 
			\begin{itemize}
				\item Is the variable suitable for summarization, visualization and checks?
				\begin{description}
					\item[Yes:] Go to stage 2. 
					\item[No:] Move on to the next variable.
				\end{description}
			\end{itemize}
		\item[Stage 2:] SVC-steps
			\begin{description}
				\item[Summarize] Call \R{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
				\item[Visualize]  Call \R{visualize()} to produce a plot visualizing the distribution of the variable. 
				\item[Check]  Call \R{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
			\end{description}
	\end{description}
\item[Output] Files for a overview document are saved to the disc and possibly also opened.
	\begin{itemize}
		\item Always a \R{rmarkdown} (.Rmd) file
		\item Possible also a html or pdf file
	\end{itemize}
\end{description}
\caption{\hl{quite odd having a figure like this? maybe do a flowchart?}}
\label{figure:cleanStructure}
\end{figure}

In \pkg{cleanR}, the function \R{clean()} is the primary workhorse and knowledge of nothing more than this tool is required to produce data cleaning outputs. The data cleaning output itself is an overview document, intended for reading by humans, in either pdf or html format. Appendix \hl{something} provides an example of a data cleaning output document, produced by calling \R{clean} on the dataset \R{toyData} available in  \R{cleanR}. The first two pages of this data cleaning output are also available in Figure~\ref{fig:example1}. \R{toyData} is a very small ($15$ by $6$), artificial dataset, whose only purpose is to illustrate the main capabilities of \R{cleanR}. The following commands load the dataset and produce the cleaning output:

@@

By default, a pdf overview document is produced, saved to the disc (in the working directory) and opened for inspection. Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts. First, an overview of what was done is presented under the title \textit{Data cleaning summary}. Secondly, each variable in the dataset is presented in turn using (up to) three tools in the \textit{Variable list}: A table summarizing key features of the variable, a figure visualizing its distribution and potentially also a list of flagged issues. For instance, in the \R{numeric}-type variable \R{var2} from \R{toyData}, \R{clean()} has identified two values that are suspected to be miscoded missing values (\R{999} and \R{NaN}), while two values were also flagged as potential outliers that should be investigated more carefully. We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable. \hl{hm, delete the last sentence? or use different variable? Only these exact two checks were performed}. \\

@@

\begin{table}
\small
\begin{tabular}{p{0.25\linewidth}p{0.45\linewidth}p{0.2\linewidth}}
\hline
Argument & Description & Default value \\
\hline

\smallskip Control \hl{...eh?}\\
\quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
\quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \R{FALSE} \\
\quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\

@@

\smallskip Control output \\
\quad \R{output} & Type of output file to be produced (html, pdf or \hl{?? what does screen mean in the documentation??}) & \R{"pdf"} \\
\quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards, and if not, should the \R{rmarkdown} file? & \R{TRUE} \\

\hline
\end{tabular}
\caption{A selection of commonly used\hl{/useful?} arguments of \R{clean()}.  \hl{Note: I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions (dealt with in next subsection)}.  \hl{Fix formatting for \R{onlyProblematic}.}}
\label{table.cleanFormals}
\end{table}

\hl{Describe arguments in Table \ref{table.cleanFormals}. Maybe do a few examples of interesting argument combinations?}

@@

\begin{table}
\centering
\begin{tabular}{p{0.35\linewidth} p{0.3\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
 p{0.01\linewidth} p{0.01\linewidth}}
  \hline
& Description &  \multicolumn{7}{c}{Variable classes} \\ \smallskip
 & &  C & F & I & L & B & N & D\\ 
  \hline \smallskip
  \textbf{\R{summaryFunction}s}  \smallskip \\
  \quad \R{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\ 
  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ &  \\ 
  \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \smallskip \\
 \textbf{\R{visualFunction}s} \smallskip \\
  \quad \R{basicVisual} & Histograms and barplots using graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \smallskip \\
 \textbf{\R{checkFunction}s} \smallskip \\
 \quad \R{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\ 
 \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & \\ 
 \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ &  \\ 
 \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\ 
 \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & \smallskip   \\ 
 \hline
\end{tabular}
\caption{\hl{Blabla, mention that C is character, F is factor, I is integer, L is labelled, B is logical (boolean), N is numeric and D is Date. \newline Also: Check that everything in here is correct (i.e. corresponds to the output of allSummaryFunctions() etc when makeXfunction has been fixed.}}
\label{table.SVCfunctions}
\end{table}

\hl{
\begin{itemize}
\item Introduce the relevant \R{clean}-arguments and the \R{defaultWhateverSummaries} etc.-functions
\item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
\item Small examples: 
\begin{itemize}
\item Add a function to one of the XXXSummaries/XXXVisuals/XXXChecks-arguments (still calling default options)
\item Remove all but a single function from one of these arguments
\item Describe what happens if the argument is NULL
\end{itemize}
\end{itemize}
}

In this section, we discuss how to control every step in \R{clean()} that actually involves a function being called on variables from the dataset. There are two stages in which this occurs, as mentioned in Figure~\ref{figure:cleanStructure}, namely:
\begin{enumerate}
\item In the precheck functions
\item In the summarize/visualize/check (SVC) step
\end{enumerate}
Each of these stages are controllable using appropriate function parameters in \code{clean}. In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs. However, if for instance the dataset at hand requires completely different visualizations, more control is needed. \pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s. \hl{Something like "the avaialable options for functions used in the precheck- and check steps are obtainable by calling \R{allCheckFunction()}, blablabla, similarly with visualFunctions and summaryFunctions. Mention Section 4 and the fact that one can expand the possibilities of e.g. allVisualFunctions() by producing new functions quite easily.}

\section{Something about the interactive cleanR mode}
While overview documents are great for documenting our work, they might not feel very natural for actually performing data cleaning or data wrangling based on the results we found. Therefore, \R{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use. Or, more precisely, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively. \hl{segway}

\hl{Maybe write here explicitly that this section introduces the three functions: \R{summarize()}, \R{check()} and \R{visualize()}?}

\subsection{Cleaning by hand: An example}
Let's say we wish to look further into a certain variable from \R{toyData}, namely \R{var2}. The data cleaning summary found some issues in this variable, and we would like to recall what these issues were. This can be done using the command
\begin{Schunk} 
\begin{Sinput} 
> check(toyData$var2)
\end{Sinput}
\begin{Soutput}
$identifyMissing
The following suspected missing value codes enter as regular values: 999, NaN.
$identifyOutliers
Note that the following possible outlier values were detected: 82, 999. $
\end{Soutput}
\end{Schunk}
\hl{delete that last \$, it's only there to make my TeX editor stop highlighting everything. Note: \$s cannot be escaped here, as Schunk is verbatim.}

@@

\begin{Schunk}
\begin{Sinput}
> missVar2 <- identifyMissing(toyData$var2)
> str(missVar2)
\end{Sinput}
\begin{Soutput}
List of 3
 $ problem      : logi TRUE
 $ message      : chr "The following suspected missing value codes enter as 
 		 regular values: \\\"999\\\", \\\"NaN\\\"."
 $ problemValues: num [1:2] 999 NaN
 - attr(*, "class")= chr "checkResult"
\end{Soutput}
\end{Schunk}
The most important thing to note here is that while the printed message is made for easy reading, the actual values of the variable causing the issue are still obtainable. If we for instance decide that the values \R{999} and \R{NaN} in \R{var2} are in fact miscoded missing values, we can easily replace them with \R{NA}s:
\begin{Schunk}
\begin{Sinput}
> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
> identifyMissing(toyData$var2}
\end{Sinput}
\begin{Soutput}
No problems found. $
\end{Soutput}
\end{Schunk}
\hl{remove that last \$}. \hl{This is quite a practical example, but it essentially also gives the basic structure for misusing cleanR by generating a loop that does autocleaning... What do you think? Perhaps those who know how to make loops also know that they shouldn't do automatic data cleaning?}

\hl{
\begin{itemize}
\item Do an example with visualize() and summarize(), like the one with check(). Especially visualize and to doEval = T thing needs a bit of special attention.
\item  Mention allCheckFunctions() etc. again here
\item Mention check(), visualize() and summarize() modes for data.frames. Maybe also advice against it, at it will often produce a lot of information at once, and such large amounts of information really should be documented.
\end{itemize}
}

@@

\subsection{Function templates}
\label{sec:functionTemplates}
\hl{metatext, mention Table \ref{table.functionTypes} again.}
